Discrete Quantum Theories and Computing

### 2018-11-19 dissertation.tex
dissertation.lyx "2.2 Quantum Probability" Updating...

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Given a finite sample space~$\Omega$ representing all possible outcomes,
and its power set $2^{\Omega}$ as the classical event space, a classical
probability measure~$\mu$ maps an event~$E$ to a number $\mu\left(E\right)\in\left[0,1\right]$
specifying how likely one of the outcome in $E$ actually happens.
To maintain the coherence, we want $\mu$ subject to the following
constraints: $\mu\left(\emptyset\right)=0$, $\mu\left(\Omega\right)=1$,
$\mu\left(\overline{E}\right)=1-\mu\left(E\right)$, and $\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

>>>>

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps an event~$E$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcome in $E$ actually happens. To maintain the
coherence, we want $\mu$ subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
and $\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

Since the previous abstraction disregards the process generating the
outcomes, this process might be a quantum experiment. Let us prepare
a beam of one kind of spin~$1$ particles. In principle, we can preform
a Stern-Gerlach type experiment to splits the beam according to the
eigenvalues of the observable $\mathbf{O}_{0}=0\,\proj{0}+1\,\proj{1}+2\,\proj{2}$
\cite{:/content/aip/journal/jmp/21/1/10.1063/1.524312,peres1995quantum}.
If we only send one particle instead of a beam, the state after the
experiment could be $\ket{0}$, $\ket{1}$, or $\ket{2}$. In another
word, the outcome of this process is $\Omega_{0}=\left\{ \ket{0},\ket{1},\ket{2}\right\} $,
and all possible events are $\emptyset$, $\left\{ \ket{0}\right\} $,
$\left\{ \ket{1}\right\} $, $\left\{ \ket{2}\right\} $, $\left\{ \ket{0},\ket{1}\right\} $,
$\left\{ \ket{0},\ket{2}\right\} $, $\left\{ \ket{1},\ket{2}\right\} $,
and $\Omega_{0}$. Sending any particular particle specifies the likelihood
of each event $\mu_{0}\left(E\right)$, and the function $\mu_{0}:2^{\Omega}\rightarrow\left[0,1\right]$
is a classical probability measure.

One special feature of quantum experiments is that the probabilities
in different experiments are correlated. Consider another experiment
similar to the previous one. This experiment send exactly the same
particle, but we use a beam splitter corresponding to the observable
$\mathbf{O}_{1}=0\,\proj{\ps}+1\,\proj{\ms}+2\,\proj{2}$, where $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$
and $\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}$. The outcome of
this process is $\Omega_{1}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $,
and all possible events are $\emptyset$, $\left\{ \ket{\ps}\right\} $,
$\left\{ \ket{\ms}\right\} $, $\left\{ \ket{2}\right\} $, $\left\{ \ket{\ps},\ket{\ms}\right\} $,
$\left\{ \ket{\ps},\ket{2}\right\} $, $\left\{ \ket{\ms},\ket{2}\right\} $,
and $\Omega_{1}$. When we formulate the likelihood of each event
as another probability measure $\mu_{1}:2^{\Omega}\rightarrow\left[0,1\right]$,
the likelihood of the same event of these two experiments are experimentally
found to be the same, i.e., $\mu_{0}\left(\emptyset\right)=\mu_{1}\left(\emptyset\right)$
and $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$.
In general, when the particle and the event are the same, so does
their probability.

Since the probability induced by the different beam splitters are
correlated, it is more natural to define the set of quantum event~$\events$
as all possible events using different beam splitters. Since we are
focus on the probability instead of outcome, and the probabilities
of events $\left\{ \ket{0},\ket{1}\right\} $ and $\left\{ \ket{\ps},\ket{\ms}\right\} $
are always the same, it is reasonable to identify them as the same
event.

, and consider the quantum probability measure $\mu:\events\rightarrow\left[0,1\right]$
as a function assigning the likelihood to every quantum event
